---
title: "PLSC 468 Pset 2"
author: "Kelly Farley, Numi Katz, and Chelsea Wang"
date: "3/11/2022"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: united
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
rm(list = ls()) # clear global environ

knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

# set working directory (change for each user)
wd <- "/Users/kellyfarley/Desktop/machine_learning/plsc468_psets"
#wd <- "/Users/numikatz/Documents/Senior_Year/Spring_22/PLSC_468/PLSC_468/plsc468_psets"
knitr::opts_knit$set(root.dir = wd)

# load libraries
library(stringr)
library(tm)
library(tidyverse)
```

# Problem 1: Supervised Learning to Classify Federalist Papers

Objective: classify 15 disputed Federalist papers as written by either Madison (H = 0) or Hamilton (H = 1)

Data: matrix "papers" with 85 "papers" and "classes;" 51 known-Hamilton, 5 known-Jay, 14 known-Madison

Process: Split 65 known Madison and Hamilton papers into training and testing sets (use discretion on proportion but consider overfitting vs validation tradeoffs); Calibrate method on training set; Evaluate method on training set (recording relevant validation measures); Classify missing papers

Output: 4x5 validation table with rows for each of the classification approaches (a) stop-words, b) non-stopwords, c) fightin' words weights, d) naive bayes) and columns for 1) overall accuracy, 2) precision for Madison, 3) precision for Hamilton, 4) recall for Madison, and 5) recall for Hamilton.

Loading papers from .rda file:

```{r}
load(file = "pset2/federalists.Rdata")
```

Cleaning, stemming, and removing rare/frequent words: (optional) (to do)

```{r}
```

## a) Mosteller and Wallace stopword dictionary approach

Calculate word weights only using stopwords in known Federalist papers

Consider appropriate document-scores threshold to classify documents and trim word weights appropriately

What is your classification?

```{r}
# using stop words from stopword() package
stops <- stopwords(kind = "en")

# note apostrophes are not removed in stops, but are in our texts
stops <- c(gsub(stops, pattern="[']", replace=''))

# make corpus
texts=VCorpus(VectorSource(papers[,1]))

# remove non stop words

dtm = as.matrix(DocumentTermMatrix(texts))
non.stops=array(TRUE,ncol(dtm))

stops=sort(stops)

col.names=colnames(dtm)
for(j in 1:length(stops)){
	ik=which(stops[j]==col.names)
	if(length(ik)>0){
		non.stops[ik]=F
	}
}

non.dtm  = dtm[,non.stops]
stop.dtm = dtm[,!non.stops]

# plot(rowSums(non.dtm),ylim=c(0,100))
# points(rowSums(stop.dtm),col='red')

# define training set

# get row IDs for known and disputed papers
papers <- as.data.frame(papers)
knownpapers_ids <- which(papers$classes == c("hamilton", "madison"), arr.ind=TRUE)
disputedpapers_ids <- which(papers$classes == "disputed", arr.ind=TRUE)

# split papers into known and disputed
knownpapers <- papers[knownpapers_ids,]
disputedpapers <- papers[disputedpapers_ids,]

# recode known papers into 0 (madison) and 1 (hamilton)
knownpapers <- knownpapers %>%
  mutate(classes = case_when(classes=="hamilton" ~ 1,
                   classes=="madison" ~ 0))

# split stop.dtm into known and disputed
stop.dtm.known <- stop.dtm[knownpapers_ids,]
stop.dtm.disputed <- stop.dtm[disputedpapers_ids,]

N=nrow(stop.dtm.known)
set.seed(1005)
s.vec=sample(1:2,replace=T,prob=c(1/2, 1/2),size=N)

# train with 1/2 of known
x.train=stop.dtm.known[which(s.vec==1),]
y.train=knownpapers$classes[which(s.vec==1)]

# test with other 1/2 of knwon
x.test=stop.dtm.known[which(s.vec==2),]
y.test=knownpapers$classes[which(s.vec==2)]

# predicted disputed
x.pred=stop.dtm.disputed
#y.pred is unknown

# a) produce a predictor score in training set

dictionW=function(x,y){
	sum.1=colSums(x[which(y==1),])
	sum.0=colSums(x[which(y==0),])

	rate.1=sum.1/(sum.0+sum.1)
	rate.0=sum.0/(sum.0+sum.1)
	var.1=rate.1*(1-rate.1)
	var.0=rate.0*(1-rate.0)

	# degenerate words get 0 weight
	rate.1[which(sum.1==0)]=0
	rate.0[which(sum.0==0)]=0
	var.1[which(sum.1==0)]=1
	var.0[which(sum.0==0)]=1

	w=(rate.1-rate.0)/(var.1+var.0)
	return(w)
}

w.train=dictionW(x=x.train,y=y.train)
top.non=sort(abs(w.train),decreasing=T)[1:20]

z.train=x.train
for(j in 1:length(w.train)){
	z.train[,j]=w.train[j]*x.train[,j]
}

Z.train=rowSums(z.train)

train.prediction=as.numeric(Z.train>mean(Z.train))

# b) testing set

z.test=x.test
for(j in 1:length(w.train)){
	z.test[,j]=w.train[j]*x.test[,j]
}

Z.test=rowSums(z.test)

test.prediction=as.numeric(Z.test>mean(Z.test))
1-mean(abs(test.prediction-y.test))

1-mean(abs(train.prediction-y.train))

# c) predicting left unknowns

z.pred=x.pred
for(j in 1:length(w.train)){
	z.pred[,j]=w.train[j]*x.pred[,j]
}

Z.pred=rowSums(z.pred)

pred.prediction=as.numeric(Z.pred>mean(Z.pred))

# e) useful to look at most predictive words

names(top.non)
```

b) Non-stopwords

Replicate part a) using non-stopwords

Calculate weights using same word-weighting approach

Consider appropriate document-scores threshold to classify documents and trim word weights appropriately

What is your classification?

c) quasi-Bayesian Fightin' Words

Use discretion over whether to include or discard stopwords or rare/frequent words

Produce Hamilton-discrimination score

What is your classification?

Which 5 words are scored the most, and which five are the least Hamiltonian?

d) Naive Bayes

Write a function to perform Naive Bayes classification to predict missing author indicators

Make sure to use Laplace smoothing

Use discretion on whether to discard stopwords or other rare/frequent words

Compare results to naiveBayes()

What are your classifications?