---
title: "PLSC 468 Final Project"
author: "Kelly Farley, Numi Katz, and Chelsea Wang"
date: "4/6/2022"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: united
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
rm(list = ls()) # clear global environ

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)

# set working directory (change for each user)
wd <- "/Users/kellyfarley/Desktop/machine_learning/plsc468_psets"
#wd <- "/Users/numikatz/Documents/Senior_Year/Spring_22/PLSC_468/PLSC_468/plsc468_psets"
knitr::opts_knit$set(root.dir = wd)

# load libraries
library(e1071)
library(stringr)
library(rjson)
library(tm)
library(tidyverse)
```

# Load Data

```{r}
raw <- fromJSON(file="/Users/kellyfarley/Desktop/oyez_pretty.json")
```

# Clean Data

# Data Structure

Dimensions
Variables
Years cases are from
Guilty / not guilty verdicts

# Most Frequent Words

Copy from Pset 1
Most frequent words across corpus, most frequent words across guilty versus not guilty, words with highest correlation

## Analysis

*What can this tell us about differences btwn guilty and not guilty verdicts?*

## Future Suggestions

# Verdict Prediction

Copy from Pset 2
Build test, training, prediction
Choose stopword, non-stopword, fighting words, Bayesian method

## Analysis

*Are we able to predict guilty vs not guilty verdicts? What's the best method for prediction?*

## Future Suggestions